Inserting an AHIR module in the NetFPGA framework

### 2011-05-29 ericsson/docs/plan.tex
documentation about putting AHIR system into NetFPGA.

The AHIR generated VHDL system is required to fit into 
a NetFPGA user-module template (described below).  Thus,
the AHIR generated VHDL system must be wrapped into
a form expected by the NetFPGA user-module template.

@@

The user module template has the form (the original
NetFPGA module is in Verilog; we use a VHDL equivalent).
\begin{verbatim}
entity module_template is
   generic (
      DATA_WIDTH: integer := 64;
      CTR_WIDTH: integer := 8;
      UDP_REG_SRC_WIDTH: integer := 2
   );
   port (
      in_data: in std_logic_vector(DATA_WIDTH-1 downto 0);
      in_ctrl: in std_logic_vector(CTRL_WIDTH-1 downto 0);
      in_wr: in std_logic;
      in_rdy: out std_logic;

      out_data: out std_logic_vector(DATA_WIDTH-1 downto 0);
      out_ctrl: out std_logic_vector(CTRL_WIDTH-1 downto 0);
      out_wr: out std_logic;
      out_rdy: in std_logic;

      clk: in std_logic;
      reset: in std_logic);
end entity;
\end{verbatim}
Thus, the NetFPGA user-module template can be viewed
as a system with a single input queue and a single output queue.

The AHIR system has form:
\begin{verbatim}
entity ahir_system is
   port(
          -- for each top-level function 
          --   (in this case int foo(int a))
          foo_a : in std_logic_vector(31 downto 0); -- argument a
          foo_ret_val_x_X: out std_logic_vector(31 downto 0); -- ret value
          foo_start: in std_logic;  -- pulse to start.
          foo_fin: out std_logic;   -- pulsed to indicate finish.

          -- for each input pipe
          --   (in this case inpipe of type int)
          inpipe_pipe_write_data: in std_logic_vector(31 downto 0); -- data
          inpipe_pipe_write_req: in std_logic_vector(0 downto 0); -- ready-in
          inpipe_pipe_write_ack: out std_logic_vector(0 downto 0); -- ready-ack.

          -- for each output pipe
          --   (in this case outpipe of type int)
          outpipe_pipe_read_data: out std_logic_vector(31 downto 0); -- data
          outpipe_pipe_read_req: in std_logic_vector(0 downto 0); -- ready-in
          outpipe_pipe_read_ack: out std_logic_vector(0 downto 0); -- ready-ack.

          clk: in std_logic;
          reset: in std_logic;
       );
end entity;
\end{verbatim}
An AHIR system can thus be viewed as having
\begin{itemize}
\item An input queue for each top-level function.
\item An output queue for each top-level function.
\item An input queue for each input pipe.
\item An output queue for each output pipe.
\end{itemize}

Thus, in order to insert an AHIR system into a NetFPGA user-module template,
all we need to do is
\begin{itemize}
\item Include an input-module which will collect data from the NetFPGA user-module
input data interface.  This input-module will feed the AHIR system input queues.
\begin{itemize}
\item This is especially simple if the AHIR system consists of a single top-level
function without pipes, since there is only one possible destination for the
incoming data.
\end{itemize}
\item Include an output-module which will collect data from the AHIR system
output queues
and forward it out of the output data interface of the NetFPGA user-module.
\begin{itemize}
\item Especially easy if there is only one top-level function in the AHIR
system without output pipes.
\end{itemize}
\item  In the general case (many top-level AHIR functions, with input and output
pipes), we will have to implement some kind of distributor/concentrator to
stream the data consistently between the NetFPGA user-module
input/output queues and the AHIR system input/output queues.
\end{itemize}

@@

There could be three approaches:
\begin{itemize}
\item The input and output modules are written as VHDL modules with
adequate parametrization.  This should work if the AHIR system is
simple enough (consists of a system with a single top-level function
without any input/output pipes, for example).  However, writing
VHDL (or Verilog) implementations is a pain for more complex scenarios.
\item The input and output modules are written as C functions and
are included into the AHIR system.  This gives much more flexibility.
However, in a complex AHIR system, this approach will limit the
ability to exploit parallelism (note: we cannot use threads in these
C functions because AhirV2 tools cannot handle threads yet).
\item The input and output modules are written as Aa modules.  This
gives not only flexibility, but also the ability to use the
parallelism in the system.
\end{itemize}

@@

What would be easiest, in my opinion would be
\begin{itemize}
\item Use the parts of Sameer's code that do the protocol
translation (between the NetFPGA user-module and an AHIR module).
\item Write the input and output modules in Aa.  For the moment,
we would do so for a specific scenario (much faster to develop and
easier to maintain).  This would remove the need to do memory 
management from the VHDL interface code (as is done in Sameer's
current implementation, in which the VHDL input module directly
accesses AHIR memory).
\item Generate the final AHIR system by including the input/output modules
written in Aa.
\item Instantiate the final AHIR system in a NetFPGA wrapper which 
just does the protocol translation between NetFPGA and AHIR.
\end{itemize}
This is indicated in Figure \ref{fig:Wrap}.   The only
additional effort is to write the input and output modules
and to write the wrap functionality in vc2vhdl.  This should
take a week.

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.7]{NetFPGAWrap.eps}
  \caption{NetFPGA Wrap Flow}
  \label{fig:Wrap}
\end{figure}

>>>>

The AHIR generated VHDL system is required to fit into 
a NetFPGA user-module template. It is relatively easy to
achieve this.

@@

The user module template has the form shown in Figure \ref{fig:NetFPGAUserModule}.
\begin{figure}
  \centering
  \includegraphics[scale=0.7]{NetFPGAUserModule.eps}
  \caption{NetFPGA User Module}
  \label{fig:NetFPGAUserModule}
\end{figure}
The interface is very simple: an input data and control bus,
and an output data and control bus.

A generic AHIR system can have the form:
\begin{figure}
  \centering
  \includegraphics[scale=0.7]{GenericAhirSystem.eps}
  \caption{Generic AHIR system}
  \label{fig:GenericAhirSystem}
\end{figure}

Thus, in order to insert a generic AHIR system into a NetFPGA user-module template:
\begin{itemize}
\item we ensure that the only interaction of the AHIR system with the outside world is
through four pipes: two input pipes for the input data and control, and
two output pipes for the output data and control.
\item The top-level modules in the AHIR system are free-running.  One solution
would be to include two top-level free-running modules which are 
responsible for interactions with the outside world (through the four pipes
discussed above).
\begin{itemize}
\item an input-module which will collect data from the input pipes and
forward to appropriate pipes/modules in the AHIR system.
\item an output-module which will collect data from internal pipes/modules
in the AHIR system and forward it to the the system output pipes.
\end{itemize}
\item The AHIR system would be instantiated in a {\em fixed} wrapper
which would be responsible for the protocol matching between the AHIR
system interface and the NetFPGA user module interface.
\end{itemize}

\section{Issues: memory management}

One issue that needs to be resolved is that of memory management.
This relates to two processes:
\begin{itemize}
\item Initialization: if the memory subsystem in the AHIR module needs
to be initialized, then this must be done explicitly by the outside
world.  The llvm2aa tool can generate memory initialization modules
that initialize the data structures in the system to their initial
values as declared in the original C program.  These modules are
auto-generated and need to be called once after the AHIR system
is reset (this call-once feature can be incorporated into vc2vhdl).
\item Exchange of information between the AHIR memory subsystem and
the outside world:  If this is needed, it would need to be explicitly 
built into the input/output modules.  Note that an exchange of
information (in general) may be initiated from ``inside'' or
from ``outside'' the AHIR system.
\end{itemize}

@@

@@

What would be easiest, in my opinion would be
\begin{itemize}
\item use the parts of Sameer's code that do the protocol
translation (between the NetFPGA user-module and an AHIR module).
\begin{itemize}
\item we implement a VHDL entity {\em NetFpgaAhirSystemWrap} which 
would instantiate an AHIR system with four interface pipes (see Figure \ref{fig:NetFpgaAhirSystemWrap}).
\begin{figure}
  \centering
  \includegraphics[scale=0.7]{NetFpgaAhirSystemWrap.eps}
  \caption{NetFpgaAhirSystemWrap Schematic}
  \label{fig:NetFpgaAhirSystemWrap}
\end{figure}
\end{itemize}
\item write the input and output modules in C/Aa and add
to the AHIR flow to generate the final system. 
\begin{itemize}
\item this can be user-written (and eventually auto-generated) for
the specific system being generated.  Note that we will have
to pay some attention to memory initialization and transfer of
information between the AHIR memory subsystem and the outside
world.
\end{itemize}
\item generate the final AHIR system by including the input/output modules
written in C/Aa.
\item Instantiate the final AHIR system in the NetFpgaAhirSystemWrap wrapper.
\end{itemize}
The process is indicated in Figure \ref{fig:Wrap}.   The only
additional effort is to write the input and output modules (in C or Aa).

\begin{figure}
  \centering
  \includegraphics[scale=0.7]{NetFPGAWrap.eps}
  \caption{NetFPGA Wrap Flow}
  \label{fig:Wrap}
\end{figure}
